---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kgp

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/kgp)](https://CRAN.R-project.org/package=kgp)
<!-- badges: end -->

This kgp data package provides metadata about populations and data about samples from the 1000 Genomes Project, including the 2,504 samples sequenced for the Phase 3 release and the expanded collection of 3,202 samples with 602 additional trios.

## Installation

You can install the released version of kgp from [CRAN](https://CRAN.R-project.org/package=kgp) with:

```r
install.packages("kgp")
```

You can install the development version of kgp from [GitHub](https://github.com/stephenturner/kgp) with:

```r
# install.packages("devtools")
devtools::install_github("stephenturner/kgp")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(kgp)
## basic example code
```


<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>. -->
