---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kgp

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/kgp)](https://CRAN.R-project.org/package=kgp)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![arXiv](https://img.shields.io/badge/arXiv-2210.00539-b31b1b.svg)](https://arxiv.org/abs/2210.00539)
<!-- badges: end -->

This kgp data package provides metadata about populations and data about samples from the 1000 Genomes Project, including the 2,504 samples sequenced for the Phase 3 release and the expanded collection of 3,202 samples with 602 additional trios.

## Installation

You can install the released version of kgp from [CRAN](https://CRAN.R-project.org/package=kgp) with:

```r
install.packages("kgp")
```

You can install the development version of kgp from [GitHub](https://github.com/stephenturner/kgp) with:

```r
# install.packages("devtools")
devtools::install_github("stephenturner/kgp")
```

## About the data

The 1000 Genomes Project data Phase 3 data contains 2,504 samples with sequence data available, and was later expanded to 3,202 samples with high coverage adding 602 trios. Data is available through the [1000 Genomes FTP site](http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/) and [GitHub](https://github.com/igsr/1000Genomes_data_indexes/). 

- Pilot publication: [An integrated map of genetic variation from 1,092 human genomes](https://www.nature.com/articles/nature11632)
- Phase 1 publication: [A map of human genome variation from population scale sequencing](https://www.nature.com/articles/nature09534)
- Phase 3 publication: [A global reference for human genetic variation](https://www.nature.com/articles/nature15393)
- Expanded high-coverage publication: [High-coverage whole-genome sequencing of the expanded 1000 Genomes Project cohort including 602 trios](https://pubmed.ncbi.nlm.nih.gov/36055201/)

There are three data sets available in the kgp package.

```{r example}
library(kgp)
data(kgp)
```

The `kgp3` data contains pedigree and population information for the 2,504 samples included in the Phase 3 release of the 1000 Genomes Project data.

```{r}
kgp3
```

The `kgpe` data contains pedigree and population information all 3,202 samples included in the expanded 1000 Genomes Project data, which includes 602 trios.

```{r}
kgpe
```

The `kgpmeta` contains population metadata for the 26 populations across five continental regions.

```{r}
kgpmeta
```

## Examples

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(kgp)
data(kgp)
```

```{r}
kgp3 %>% 
  count(region)
```

```{r}
kgp3 %>% 
  count(region, population)
```

```{r kgp3barplot, fig.width=9, fig.height=12}
kgp3 %>% 
  count(region, population) %>% 
  arrange(region, n) %>% 
  mutate(population=forcats::fct_inorder(population)) %>% 
  ggplot(aes(population, n)) + 
  geom_col(aes(fill=region)) + 
  labs(fill=NULL, x=NULL, x="N") + 
  coord_flip() + 
  theme_bw() + 
  theme(legend.position="bottom")
```
